Mixed effect tweedie model in TMB

### 2019-04-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
exploring effect of daily SST on bottom temperature

plan(multiprocess)
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
glimpse(d_trawl)
nrow(d_trawl)
head(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3
)
unique(d_trawl$year)
d_trawl[is.na(d_trawl$year), ]$year <- 2003
```

@@

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
write.csv(cv$summary, file = "cv-test.csv")
cv_test <- read_csv(file = "cv-test.csv")
glimpse(cv)
View(cv$summary)
# plot_cv_meshes(cv$cv_model_set[[1]])
# plot_cv_meshes(cv$cv_model_set[[2]])
```

@@

non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
View(non_time_varying$summary)
write.csv(non_time_varying$summary, file = "cv-non-time-varying.csv")
saveRDS(non_time_varying, file = "cv-non-time-varying.rds")
glimpse(non_time_varying$summary)
```

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and non showed a linear effect of depth
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

@@

View(time_varying$summary)
write.csv(time_varying$summary, file = "cv-time-varying.csv")
saveRDS(time_varying, file = "cv-time-varying.rds")
glimpse(time_varying$summary)
```

## Now vary random effect configuation for top model(s) 
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

vary_rf_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(vary_rf_knots$summary)
write.csv(vary_rf_knots$summary, file = "cv-varying-rf.csv")
saveRDS(vary_rf_knots, file = "cv-varying-rf.rds")
glimpse(vary_rf_knots$summary)
```

## Change seeds 
```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(TRUE),
  n_knots = c(120),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

View(more_seeds$summary)
write.csv(more_seeds$summary, file = "cv-more-seeds.csv")
saveRDS(more_seeds, file = "cv-more-seeds.rds")
glimpse(more_seeds$summary)
```

Load saved batches:
```{r, eval=FALSE}
# non_time_varying <- readRDS(file = "cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-refine-knots.rds")

non_time_varying <- read.csv(file = "cv-non-time-varying.csv")
time_varying <- read.csv(file = "cv-time-varying.csv")
vary_rf <- read.csv(file = "cv-varying-rf.csv")
non_time_varying1 <- read.csv(file = "cv-non-time-varying1.csv")
time_varying1 <- read.csv(file = "cv-time-varying1.csv")
vary_rf1 <- read.csv(file = "cv-varying-rf1.csv")
refine_knots1 <- read.csv(file = "cv-refine-knots1.csv")
more_seeds <- read.csv(file = "cv-more-seeds.csv")
more_seeds1 <- read.csv(file = "cv-more-seeds1.csv")

@@

Current top model
```{r}
plan(multiprocess)
temp3_rw <- sdmTMB_cv(
  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
  formula = temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", family = gaussian(link = "identity"),
  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
)
# temp3_rw$models
sum(temp3_rw$sum_loglik)
```

@@

```{r}
m_temp3_rw <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

predictions3_rw <- predict(m_temp3_rw)
predictions3_rw$data$residuals <- residuals(m_temp3_rw)
cor(predictions3_rw$data$est, predictions3_rw$data$temperature_c)^2 #
# aic(m_temp3_rw)
```

Residual plots:
```{r}
ggplot(predictions3_rw$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

>>>>

plan(multiprocess)
```

Load data
```{r}
# d_trawl1 <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
# nrow(d_trawl1)
d_trawl <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl-SST.rds"))
nrow(d_trawl)
# d_trawl$log_temperature_c <- log(d_trawl$temperature_c)
unique(d_trawl$year)
d_trawl[is.na(d_trawl$year), ]$year <- 2003
d_trawl <- d_trawl[!is.na(d_trawl$depth_m), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl <- mutate(d_trawl,
  depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(SST),
  depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth_m>50,0,1)
)
glimpse(d_trawl)
```

Explore SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(dplyr::filter(d_trawl, depth_m <50), aes(SST,temperature_c, colour = depth_m)) + geom_point()
```

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3_nosst <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

m_temp3
m_temp3_nosst
```

```{r}

predictions3 <- predict(m_temp3)
predictions3b <- predict(m_temp3_nosst)

predictions3$data$residuals <- residuals(m_temp3)
predictions3b$data$residuals <- residuals(m_temp3_nosst)

nd <- data.frame(sst_resid = predictions3$data$residuals, nosst_resid = predictions3b$data$residuals, shallow = d_trawl$shallow, depth = d_trawl$depth_m)

ggplot(filter(nd, shallow == 1), aes(nosst_resid,sst_resid, colour = as.factor(shallow))) + 
  geom_point() + geom_abline(intercept = 0, slope = 1)
```

@@

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
write.csv(cv$summary, file = "cv-output/cv-test.csv")
cv_test <- read_csv(file = "cv-output/cv-test.csv")
View(cv$summary)
# plot_cv_meshes(cv$cv_model_set[[1]])
# plot_cv_meshes(cv$cv_model_set[[2]])
```

@@

non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
View(non_time_varying$summary)
write.csv(non_time_varying$summary, file = "cv-output/cv-non-time-varying.csv")
saveRDS(non_time_varying, file = "cv-output/cv-non-time-varying.rds")
glimpse(non_time_varying$summary)
```

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and none showed a linear effect of depth
All are better than fixed slope models
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

@@

View(time_varying$summary)
write.csv(time_varying$summary, file = "cv-output/cv-time-varying.csv")
saveRDS(time_varying, file = "cv-output/cv-time-varying.rds")
glimpse(time_varying$summary)
```

## Now vary random effect configuation for top model(s) 
a separate spatial field is in all models with highest likelihoods for each seed
AR1 doesn't make much difference (differences between years appear independent)
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

vary_rf <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(vary_rf$summary)
write.csv(vary_rf$summary, file = "cv-output/cv-varying-rf.csv")
saveRDS(vary_rf, file = "cv-output/cv-varying-rf.rds")
glimpse(vary_rf$summary)
```

## Change knot number

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(TRUE),
  n_knots = c(80,120,140),
  seed = c(1111, 2222, 3333, 4444)
)

more_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(more_knots$summary)
write.csv(more_knots$summary, file = "cv-output/cv-more-seeds.csv")
saveRDS(more_knots, file = "cv-output/cv-more-seeds.rds")
glimpse(more_knots$summary)
```

## Change seeds 
Also increased knot number too 
Model without SST is always very slightly higher ll

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(TRUE),
  n_knots = c(130),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

@@

View(more_seeds$summary)
write.csv(more_seeds$summary, file = "cv-output/cv-more-seeds.csv")
saveRDS(more_seeds, file = "cv-output/cv-more-seeds.rds")
glimpse(more_seeds$summary)
```

Load saved batches:
```{r, eval=FALSE}
# non_time_varying <- readRDS(file = "cv-output/cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-output/cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-output/cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-output/cv-refine-knots.rds")

non_time_varying <- read.csv(file = "cv-output/cv-non-time-varying.csv")
time_varying <- read.csv(file = "cv-output/cv-time-varying.csv")
vary_rf <- read.csv(file = "cv-output/cv-varying-rf.csv")
non_time_varying1 <- read.csv(file = "cv-output/cv-non-time-varying1.csv")
time_varying1 <- read.csv(file = "cv-output/cv-time-varying1.csv")
vary_rf1 <- read.csv(file = "cv-output/cv-varying-rf1.csv")
refine_knots1 <- read.csv(file = "cv-output/cv-refine-knots1.csv")
more_seeds <- read.csv(file = "cv-output/cv-more-seeds.csv")
more_seeds1 <- read.csv(file = "cv-output/cv-more-seeds1.csv")

@@

Current top model
```{r}
plan(multiprocess)
temp_rw <- sdmTMB_cv(
  data = d_trawl, n_knots = 130, x = "X", y = "Y", k_folds = 10,
  formula = temperature_c ~ 0 + as.factor(year) + SST,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", family = gaussian(link = "identity"),
  ar1_fields = FALSE, include_spatial = TRUE, seed = 999
)
sum(temp_rw$sum_loglik)
```

@@

```{r}
m_temp_rw <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

predictions_rw <- predict(m_temp_rw)
predictions_rw$data$residuals <- residuals(m_temp_rw)
cor(predictions_rw$data$est, predictions_rw$data$temperature_c)^2 #
# aic(m_temp3_rw)
```

Residual plots:
```{r}
ggplot(predictions_rw$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions_rw$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

Does a fixed interaction with SST do any better?

```{r}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST*depth_scaled + SST*depth_scaled2 + SST*depth_scaled3,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

predictions3 <- predict(m_temp3)
predictions3$data$residuals <- residuals(m_temp3)
cor(predictions3$data$est, predictions3$data$temperature_c)^2 #
```

```{r}
ggplot(predictions3$data, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions3$data, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```
